---
title: "Analyze intrataxonomic distances."
author: "Thelma Panaïotis"
format: 
  html:
    toc: true
    embed-resources: true
editor: visual
lightbox: true
execute:
  warning: false
  cache: true
  freeze: false
---

```{r set_up}
#| cache: false
source("utils.R")
```

## Read saved distances

List saved distances and ream them

```{r load}
## Intra distances
# list processed files
processed <- list.files("data", pattern = "04.", full.names = TRUE)

# load data
res <- sapply(processed, function(x) mget(load(x)), simplify = TRUE)

# get summary data (1st line)
df_intra <- res[1,] %>% bind_rows()
# get distances (2nd line)
df_intra_dist <- res[2,] %>% bind_rows()

## Null distances
load("data/02a.f_val_dist.Rdata")
load("data/02b.rq_coef.Rdata")

## Apply log transformation for plotting
f_val_dist <- f_val_dist %>% 
  mutate(
    log_n_dist = log10(n_dist),
    log_test_stat = log10(test_stat)
  )

df_intra <- df_intra %>% 
    mutate(
    log_n_dist = log10(n_dist),
    log_test_stat = log10(test_stat)
  )
```

## Analyse Kuiper statistic

Plot test stat VS number of computed distances in a log-transformed space.

```{r f_vs_n_dist}
#| fig-column: body-outset
#| out-width: 100%

## Generate data to plot a ribbon between the regression lines
# limits for n_dist
lim_dist <- c(
  floor_dec(min(df_intra$n_dist), level = -(nchar(paste(min(df_intra$n_dist))) - 1)),
  ceiling_dec(max(df_intra$n_dist), level = -(nchar(paste(max(df_intra$n_dist))) - 1))
)
# Generate ribbon
rib_data <- tibble(n_dist = lim_dist) %>% 
  mutate(
    # apply log-transformation
    log_n_dist = log10(n_dist),
    # compute estimated kuiper-stat from slope and intercept
    ymin = rq_coef %>% filter(tau == 0.05) %>% pull(slope) * log_n_dist + rq_coef %>% filter(tau == 0.05) %>% pull(intercept),
    ymax = rq_coef %>% filter(tau == 0.95) %>% pull(slope) * log_n_dist + rq_coef %>% filter(tau == 0.95) %>% pull(intercept)
  ) %>% 
  # reformat and reorder to plot a polygon
  pivot_longer(ymin:ymax, values_to = "y") %>% 
  mutate(order = c(1, 2, 4, 3)) %>% 
  arrange(order)

## Plot results
ggplot() +
  geom_boxplot(data = f_val_dist, aes(x = log_n_dist, y = log_test_stat, group = log_n_dist), colour = "gray") +
  geom_polygon(data = rib_data, aes(x = log_n_dist, y = y), alpha = 0.1) +
  geom_point(data = df_intra, aes(x = log_n_dist, y = log_test_stat, fill = taxon, colour = taxon, shape = taxon)) +
  scale_fill_paletteer_d(`"khroma::discreterainbow"`) +
  scale_colour_paletteer_d(`"khroma::discreterainbow"`) +
  scale_shape_manual(values = rep(21:25, 6)) +
  scale_x_continuous(labels = label_math(expr = 10^.x, format = force)) +
  scale_y_continuous(labels = label_math(expr = 10^.x, format = force)) +
  labs(x = "N distances", y = "Kuiper statistis", colour = "Taxon", fill = "Taxon", shape = "Taxon") 
```

Four groups seem to have distances that differ from expected under a random distribution:

-   Acantharea

-   Cop_small

-   Cop_Calanoida

-   Cop_other

Let’s plot their distances.

```{r all_dist}
#| fig-column: screen-inset
#| out-width: 100%
#| fig-width: 12
#| fig-height: 5
df_intra_dist %>% 
  filter(taxon %in% c("Acantharea", "Cop_small", "Cop_Calanoida", "Cop_other")) %>% 
  pivot_longer(dist:dist_rand) %>% 
  ggplot() +
  geom_density(aes(x = value, colour = name)) +
  facet_wrap(~taxon, scales = "free")
```

And plot their ECDF.

```{r all_ecdf}
#| fig-column: screen-inset
#| out-width: 100%
#| fig-width: 12
#| fig-height: 5
df_intra_dist %>% 
  filter(taxon %in% c("Acantharea", "Cop_small", "Cop_Calanoida", "Cop_other")) %>% 
  pivot_longer(dist:dist_rand) %>% 
  ggplot() +
  stat_ecdf(aes(x = value, colour = name)) +
  facet_wrap(~taxon, scales = "free")
```

For the 4  groups, distances distribution is shifted to the left compared to null distances: individuals are closer to each other than expected under a random distribution.
