---
title: "Analyze intrataxonomic distances, using small distances only."
author: "Thelma Panaïotis"
format: 
  html:
    toc: true
    embed-resources: true
editor: visual
lightbox: true
execute:
  warning: false
  cache: true
  freeze: false
---

```{r set_up}
#| echo: false
#| cache: false
source("utils.R")
```

## Load data

```{r read}
#| cache.lazy: false
load("data/02a.f_val_dist_small.Rdata")
load("data/02b.rq_coef_small.Rdata")

## Intra distances
# list processed files
processed <- list.files("data", pattern = "04\\.", full.names = TRUE)

# load data
res <- sapply(processed, function(x) mget(load(x)), simplify = TRUE)

# get summary data (1st line)
df_intra <- res[1,] %>% bind_rows()
# get distances (2nd line)
df_intra_dist <- res[2,] %>% bind_rows()

## Assign colour and shape to each taxon
df_intra <- df_intra %>% 
  mutate(
    colour = as.character(paletteer_d(`"khroma::discreterainbow"`, n = 28)),
    shape = rep(21:25, 6)[1:28]
  )
```

## Process data

### Null distances

```{r prep_null}
# Apply log transformation for plotting
f_val_dist <- f_val_dist %>% 
  mutate(
    log_n_dist = log10(n_dist),
    log_test_stat = log10(test_stat)
  )

## Generate data to plot a ribbon between the regression lines
# limits for n_dist
lim_dist <- c(50, 2e9)
# Generate ribbon
rib_data <- tibble(n_dist = lim_dist) %>% 
  mutate(
    # apply log-transformation
    log_n_dist = log10(n_dist),
    # compute estimated kuiper-stat from slope and intercept
    ymin = rq_coef %>% filter(tau == 0.05) %>% pull(slope) * log_n_dist + rq_coef %>% filter(tau == 0.05) %>% pull(intercept),
    ymax = rq_coef %>% filter(tau == 0.95) %>% pull(slope) * log_n_dist + rq_coef %>% filter(tau == 0.95) %>% pull(intercept)
  ) %>% 
  # reformat and reorder to plot a polygon
  pivot_longer(ymin:ymax, values_to = "y") %>% 
  mutate(order = c(1, 2, 4, 3)) %>% 
  arrange(order)
```

### Plankton distances

We need to filter distances to keep only the small distances, and perform the Kuiper test using these filtered distances.

```{r filter_dist}
# Keep only small distances
df_intra_dist_small <- df_intra_dist %>% 
  pivot_longer(dist:dist_rand) %>% 
  mutate(value = value * 51 / 10000) %>% # from pixel to cm
  filter(value < dist_thr_cm)

# Perform Kuiper test using only small distances
df_intra_small <- mclapply(df_intra$taxon, function (ta) {
  # Get distances
  d1 <- df_intra_dist_small %>% filter(taxon == ta) %>% filter(name == "dist") %>% pull(value) # plankton
  d2 <- df_intra_dist_small %>% filter(taxon == ta) %>% filter(name == "dist_rand") %>% pull(value) # null
  
  # Number of distances
  n_dist_small <- (length(d1) + length(d2)) / 2
  
  # Kuiper test
  kt <- kuiper_test(d1, d2)
  
  # Return results
  return(
    tibble(
    taxon = ta,
    prop_dist_small = n_dist_small / 10000, # proportion of small distances based on the quantiles
    test_stat = kt[1],
    p_value = kt[2]
  ))
  
}, mc.cores = n_cores) %>% 
  bind_rows()

# compute number of small distances based on proportion of small distances and the total number of distances
df_intra_small <- df_intra %>% 
  select(taxon, n_dist, colour, shape) %>% 
  left_join(df_intra_small, by = join_by(taxon)) %>% 
  mutate(
    n_dist_small = n_dist * prop_dist_small,
    log_n_dist_small = log10(n_dist),
    log_test_stat = log10((test_stat))
  )

# Detect taxa which are above the ribbon and have the minimum number of distances required
df_intra_small <- df_intra_small %>% 
  mutate(
    above = log_test_stat > log_n_dist_small * (rq_coef %>% filter(tau == 0.95) %>% pull(slope)) + (rq_coef %>% filter(tau == 0.95) %>% pull(intercept)), # above the polygon
    keep = n_dist_small > n_dist_min, # enough distances
    above = above & keep # combination of both
    )
df_intra_small_above <- df_intra_small %>% filter(above) # needed to keep the same colour and shape across plots
```

## Results

### Kuiper statistic VS number of distances

Plot Kuiper statistic VS number of distances.

```{r kuiper_plot}
#| fig-column: body-outset
#| out-width: 100%
# With all taxa
ggplot() + 
  geom_boxplot(data = f_val_dist, aes(x = log_n_dist, y = log_test_stat, group = log_n_dist), colour = "gray") +
  geom_polygon(data = rib_data, aes(x = log_n_dist, y = y), alpha = 0.1) +
  geom_point(data = df_intra_small, aes(x = log_n_dist_small, y = log_test_stat, colour = taxon, fill = taxon, shape = taxon)) +
  scale_fill_manual(values = df_intra_small$colour) +
  scale_colour_manual(values = df_intra_small$colour) +
  scale_shape_manual(values = df_intra_small$shape) +
  scale_x_continuous(labels = label_math(expr = 10^.x, format = force)) +
  scale_y_continuous(labels = label_math(expr = 10^.x, format = force)) +
  labs(x = "N distances", y = "Kuiper statistic", colour = "Taxon", fill = "Taxon", shape = "Taxon") 

# With only taxa above ribbon
ggplot() +
  geom_boxplot(data = f_val_dist, aes(x = log_n_dist, y = log_test_stat, group = log_n_dist), colour = "gray", outlier.shape = NA) +
  geom_polygon(data = rib_data, aes(x = log_n_dist, y = y), alpha = 0.1) +
  geom_point(data = df_intra_small %>% filter(above), aes(x = log_n_dist_small, y = log_test_stat, fill = taxon, colour = taxon, shape = taxon)) +
  geom_point(data = df_intra_small %>% filter(!above), aes(x = log_n_dist_small, y = log_test_stat), colour = "gray") +
  scale_fill_manual(values = df_intra_small_above$colour) +
  scale_colour_manual(values = df_intra_small_above$colour) +
  scale_shape_manual(values = df_intra_small_above$shape) +
  scale_x_continuous(labels = label_math(expr = 10^.x, format = force), breaks = seq(2, 8, by = 2)) +
  scale_y_continuous(labels = label_math(expr = 10^.x, format = force)) +
  labs(x = "N distances", y = "Kuiper statistic", colour = "Taxon", fill = "Taxon", shape = "Taxon") 
```

### Distribution of distances

Plot distribution of distances.

```{r dist}
#| fig-column: body-outset
#| out-width: 100%
# Plot density
df_intra_dist_small %>% 
  filter(taxon %in% df_intra_small_above$taxon) %>% 
  ggplot() +
  geom_density(aes(x = value, colour = name)) +
  scale_colour_manual(
    values = c("dist_rand" = "grey20", "dist" = "red"),
    labels = c("plankton", "null")
    ) +
  labs(x = "Distance (cm)", y = "Density", colour = "Type") +
  facet_wrap(~taxon, scales = "free")
```

Plot ECDF.

```{r ecdf}
#| fig-column: body-outset
#| out-width: 100%
# Plot ECDF
df_intra_dist_small %>% 
  filter(taxon %in% df_intra_small_above$taxon) %>% 
  ggplot() +
  stat_ecdf(aes(x = value, colour = name)) +
  scale_colour_manual(
    values = c("dist_rand" = "grey20", "dist" = "red"),
    labels = c("plankton", "null")
    ) +
  facet_wrap(~taxon, scales = "free")
```

### Interaction strength & confidence

```{r z_scores}
#| fig-column: body-outset
#| out-width: 100%
#| fig-width: 8
#| fig-height: 5

df_intra_scores <- df_intra_small %>% 
  mutate(
    # compute estimated mean, 5th and 95th percentiles kuiper-stat from slope and intercept
    ymin = rq_coef %>% filter(tau == 0.05) %>% pull(slope) * log_n_dist_small + rq_coef %>% filter(tau == 0.05) %>% pull(intercept),
    ymean = rq_coef %>% filter(tau == "mean") %>% pull(slope) * log_n_dist_small + rq_coef %>% filter(tau == "mean") %>% pull(intercept),
    ymax = rq_coef %>% filter(tau == 0.95) %>% pull(slope) * log_n_dist_small + rq_coef %>% filter(tau == 0.95) %>% pull(intercept),
    # compute sigma
    sigma = (ymax - ymin) / (2 * 1.645),
    # compute z-score
    z_score = (log_test_stat - ymean)/sigma
  )

ggplot() +
  geom_boxplot(data = f_val_dist, aes(x = log_n_dist, y = log_test_stat, group = log_n_dist), colour = "gray", outlier.shape = NA) +
  geom_polygon(data = rib_data, aes(x = log_n_dist, y = y), alpha = 0.1) +
  geom_point(data = df_intra_scores %>% filter(above), aes(x = log_n_dist_small, y = log_test_stat, colour = z_score)) +
  scale_colour_viridis_c() +
  scale_x_continuous(labels = label_math(expr = 10^.x, format = force)) +
  scale_y_continuous(labels = label_math(expr = 10^.x, format = force)) +
  labs(x = "N distances", y = "Kuiper statistic", colour = "Z-score") 
```

Plot Z-score VS Kuiper statistic.

```{r plot_z}
#| fig-column: body-outset
#| out-width: 100%
#| fig-width: 8
#| fig-height: 5
df_intra_scores %>% 
  filter(above) %>% 
  #filter(!str_detect(taxon, "Collodaria_colonial")) %>% 
  ggplot() + 
  geom_point(aes(x = test_stat, y = z_score, colour = taxon, fill = taxon, shape = taxon)) +
  scale_colour_manual(values = df_intra_small_above$colour) +
  scale_fill_manual(values = df_intra_small_above$colour) +
  scale_shape_manual(values = df_intra_small_above$shape) +
  labs(x = "Kuiper statistic", y = "Z-score", colour = "Taxon", fill = "Taxon", shape = "Taxon") +
  expand_limits(x = 0, y = 0) +
  xlim(c(0, 0.05))
```

::: {.callout-note icon="false"}
## Conclusion

**When focusing on small distances only, we detect more taxa for which the distribution of distances differs from those expected under a random distribution of organisms.**
:::

## Save

Save Z-score and Kuiper stat.

```{r save}
save(df_intra_scores, file = "data/04c.df_intra_scores_small.Rdata")
```
